Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-11 ms.tex
fixed 6.3

It is also possible that many of the nuclei from the \citet{Miller+15}
sample are not accreting in a steady fashion.  The ADAF efficiency
assumed in the right panel of Figure \ref{fig:miller} in some ways
represents the maximum value, since it does not take into account
outflows from the accretion flow that would suppress the accreiton
rate reaching large radii.  Nevertheless, many lower black hole masses
of $M_{\bullet} \lesssim 10^{8}M_{\odot}$ show X-ray luminosities lie
well above the predicted average value for stable accretion (black
line), suggesting that these galaxies either experience lower levels
of heating than average galaxies of their mass, or that they are not
accreting stably at all ($\S\ref{sec:cycle}$).  Indeed, for the lowest
mass holes, $M_{\bullet} \lesssim 10^{8}M_{\odot}$, $L_{\rm X}$
exceeds the one corresponding to the maximum thermally-stable
accretion rate, $\dot{M}_{\rm TI}$ (yellow line).

Figure \ref{fig:bh_growth} (top panel) shows that across all black
hole masses, Compton heating alone cannot prevent thermal
instabilities from developing in a steady-state accretion flow
($\dot{M}_{\rm C} > \dot{M}_{\rm TI}$).  Thus, if the stellar
population is sufficiently old (or the star formation sufficiently
intermittent) then nuclear gas could undergo cyclic oscillations of
thermal instability and high accretion rates, followed by quiescent
periods once gas has been consumed and/or AGN feedback becomes
effective.  Such cycles are proposed to occur on large scales and long
timescales in elliptical galaxies (\citealt{Ciotti+10}), an idea
supported, e.g. , by GALEX observations showing that a small fraction
of early-type galaxies in the local universe have undergone recent ($<
1-2$ Gyr old) star formation episodes (\citealt{Donas+07}).  If such
cyclic behavior also occurs on small, parsec scales (e.g.) then the
large scatter in the accretion rates and hence X-ray luminosities
shown in Fig \ref{fig:miller} could also be understood by the wide
range of gas masses experienced over the course of a cyclic episode
between instabilities and periodic nuclear outbursts
(e.g.~\citealt{Ciotti+10}).  In support of this possibility, we find
that core galaxies are more likely to possess higher accretion rates
and potentially be thermally unstable.  Past studies have found that
core galaxies show a wide range of nuclear radio and X-ray
luminosities (suggesting cyclic accretion episodes), while cusp
galaxies are confined below a threshold (e.g., \citealt{Bender+89,
  Pellegrini99, Capetti&Balmaverde05}), consistent with
thermally-stable, steady-state picture.

The cyclic model predicts that high $L_{X}$ nuclei should be more
common in galaxies with younger stellar ages, a result which appears
in tension with the finding by \citet{Pellegrini10} that galaxies with
younger stellar populations are observed to host less X-ray luminous
nuclei, at least for stellar ages $<3$ Gyr.  An alternative
explanation for this observation within a thermally stable,
steady-state accretion picture is that young stars provide higher
stellar wind heating, reducing the mass accretion rate and hence
nuclear X-ray luminosity.  {\bf AG Challenge: \citet{HeckmanBest:2014a}
  find that ${\rm L_{[OIII]}}$ (a proxy for the bolometric luminosity)
  ${\rm L_{[OIII]}/M_\star}$ (a proxy for the Eddington ratio)
  increases with decreasing 4000 $\AA$ break index (decreasing stellar
  age) (see their Figure 14). This suggests that the accretion rate is
  actually increasing towards younger stellar ages. At young ages
  there actually seems to be plenty of gas around, which may be
  obscuring the x-rays and causing an anamolously low measurement of
  $L_X$.

  For a starburst the black hole accretion rate lags the star
  formation rate by 200 Myr (\citet{WildHeckman+:2010a} Figure
  9). This could be due to SNe feed back suppressing the black hole
  accretion rate (massive star winds do the same thing by themselves,
  but they cannot be sustained for 200 Myr. 
}

@@

The bottom panel of Figure~\ref{fig:bh_growth} shows the SMBH growth
time, $t_{\rm grow} = M_{\bullet}/\dot{M}$, for each accretion rate
shown in the top panel.  For black holes accreting steadily given the
stellar wind heating associated with the average star formation
history of their galaxies, we see that $t_{\rm grow}$ exceeds the
Hubble time by 1-3 orders of magnitude across all black hole masses.
Steady accretion therefore cannot explain the growth of low mass black
holes, a fact which is not surprising given that approximately half of
this growth occurs in AGN radiating within 10 per cent of their
Eddington rate (\citealt{Heckman+04}).  Such high accretion rates
likely instead require an source of gas external to the nuclear
region, triggered either by galaxy mergers associated with the
hierachical growth of structure or thermal instabilities on larger,
galactic scales (e.g.~\citealt{Ciotti+10}).

>>>>

It is also possible that many of the galactic nuclei from the \citet{Miller+15} sample are not accreting in a steady fashion.  The efficiency assumed in the right panel of Figure \ref{fig:miller} in some ways represents a maximum value, since it does not take into suppression of the accretion rate due to disk outflows on small scales.  If the actual X-ray efficiency is lower than assumed, then many of the detections in the case of low SMBH masses, $M_{\bullet} \lesssim 10^{8}M_{\odot}$, would lie even further above above the predicted average value for stable accretion (black line).  These galaxies either experience significantly lower levels of heating than average galaxies of their mass, or that they are not accreting stably at all.  Indeed, for the lowest mass holes, $M_{\bullet} \lesssim 10^{8}M_{\odot}$, $L_{\rm X}$ nearly exceeds that predicted for the maximum thermally-stable accretion rate, $\dot{M}_{\rm TI}$ (yellow line).  It is not surprising that this sample is more likely to be thermally unstable as compared to an average galaxy with this black hole mass, because the \citet{Miller+15} sample includes only early-type galaxies with on averager older stellar populations.

Figure \ref{fig:bh_growth} (top panel) also shows that across all black hole masses, Compton heating also cannot prevent average accretion rates which are thermally unstable, i.e. $\dot{M}_{\rm C} > \dot{M}_{\rm TI}$.  In the case of thermal instabilities, nuclear gas may undergo cyclic oscillations of rapid cooling and high accretion rates, followed by quiescent
periods once gas has been consumed and/or AGN feedback becomes effective.  Such cycles are proposed to occur on large scales and long timescales in elliptical galaxies (\citealt{Ciotti+10}), an idea possibly supported by, e.g., GALEX observations showing that a small fraction of early-type galaxies in the local universe have undergone recent ($< 1-2$ Gyr old) star formation episodes (\citealt{Donas+07}).  Similiar cyclic AGN activity have been proposed to occur on the smaller scale of the sphere of influence (e.g.~\citealt{Yuan&Li11}, \citealt{Cuarda+15}).

If such cyclic outbursts also manifest on the properties of gas on small, parsec scales, then the large scatter in the accretion rates and hence X-ray luminosities in Fig \ref{fig:miller} could result from the wide range of inflow rates experienced over the course of a cyclic episode between instabilities and periodic nuclear outbursts (e.g.~\citealt{Ciotti+10}).  Due to their deeper potential well, we find that core galaxies are more likely to possess higher accretion rates and potentially be thermally unstable for a fixed wind heating rate than cusp galaxies (Fig.~XXX).  This could help explain why core galaxies show a wide range of nuclear radio and X-ray luminosities (suggesting cyclic accretion episodes), while cusp galaxies are confined below a threshold (e.g., \citealt{Bender+89,   Pellegrini99, Capetti&Balmaverde05}), consistent with thermally-stable, steady-state picture.

If the galaxies with measured $L_{\rm X}$ at low $M_{\bullet}$ in Fig \ref{fig:miller} represent thermally-unstable galaxies, then many of the upper limits may correspond to a separate ``thermally stable" population, for instance including SgrA*.  The possibility of a mixed population of stable versus unstable galaxies calls into question the use of simple extrapolations of power-law fits to the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ to constrain the occupation fraction of black holes in low mass galaxies (\citealt{Miller+15}).

%The cyclic model predicts that high $L_{X}$ nuclei should be more
%common in galaxies with younger stellar ages, a result which appears
%in tension with the finding by \citet{Pellegrini10} that galaxies with
%younger stellar populations are observed to host less X-ray luminous
%nuclei, at least for stellar ages $<3$ Gyr.  An alternative
%explanation for this observation within a thermally stable,
%steady-state accretion picture is that young stars provide higher
%stellar wind heating, reducing the mass accretion rate and hence
%nuclear X-ray luminosity.  {\bf AG Challenge: \citet{HeckmanBest:2014a}
 % find that ${\rm L_{[OIII]}}$ (a proxy for the bolometric luminosity)
  %${\rm L_{[OIII]}/M_\star}$ (a proxy for the Eddington ratio)
  %increases with decreasing 4000 $\AA$ break index (decreasing stellar
 % age) (see their Figure 14). This suggests that the accretion rate is
  %actually increasing towards younger stellar ages. At young ages
  %there actually seems to be plenty of gas around, which may be
  %obscuring the x-rays and causing an anamolously low measurement of
  %$L_X$.

  %For a starburst the black hole accretion rate lags the star
  %formation rate by 200 Myr (\citet{WildHeckman+:2010a} Figure
 % 9). This could be due to SNe feed back suppressing the black hole
  %accretion rate (massive star winds do the same thing by themselves,
 % but they cannot be sustained for 200 Myr. 
%}

@@

The bottom panel of Figure~\ref{fig:bh_growth} shows the SMBH growth
time, $t_{\rm grow} = M_{\bullet}/\dot{M}$, for each accretion rate
shown in the top panel.  For black holes accreting steadily given the
stellar wind heating associated with the average star formation
history of their galaxies, we see that $t_{\rm grow}$ exceeds the
Hubble time by 1$-$3 orders of magnitude across all black hole masses.
Steady accretion therefore cannot explain the growth of low mass black
holes, a fact which is not surprising given that approximately half of
this growth occurs in AGN radiating within 10 per cent of their
Eddington rate (\citealt{Heckman+04}).  Such high accretion rates
likely instead require an source of gas external to the nuclear
region, triggered either by galaxy mergers associated with the
hierachical growth of structure or thermal instabilities on larger,
galactic scales (e.g.~\citealt{Ciotti+10}).
